---
title: "R_tutorial"
output:
  rmdformats::downcute
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

We will be using [R version 4.0.0](https://www.r-project.org) for this tutorial, feel free to copy and mess with the code snippets to help your understanding.

I recommend you to download and install [RStudio](https://rstudio.com) to help ypu visualize and edit your codes.

# Basic concepts

## Numbers

You can just type any number and R will give you the same number

```{r}
1
```
## Strings

In order to show a string you must use 'single' or "double" quotes

```{r}
"HI"

'Hello'
```


## Arithmetic

You can type the equation and R will compute your answer

```{r}
2/4
```
```{r, echo=FALSE}
options(scipen=999)
Name = c('Addition','Subtraction','Multiplication','Division','Exponent','Modulus','Integer Division')
Operator = c("+","-","*",'/','^','%%','%/%') 
Example = c('10+20','10-20','10*20','10/20','10^20','10%%20','10%/%20') 
Result = c(10+20,10-20,10*20,10/20,10^20,10%%20,10%/%20)
df = data.frame(Name,Operator,Example,Result)
knitr::kable(df, caption = 'Cheat sheet fpr arithmetic operations in R')
```

## Booleans 

Binary values, either `r TRUE` or `r FALSE`

```{r, echo = FALSE}
options(scipen=999)
Name = c('Equal','Not equal','Less than','Greater than','Less than or equal to','Greater than or equal to')
Operator = c('==','!=','<','>','<=','>=') 
Example = c('10==20','10!=20','10>20','10<20','10<=20','10>=20') 
Result = c(10==20,10!=20,10>20,10<20,10<=20,10>=20)
df = data.frame(Name,Operator,Example,Result)
knitr::kable(df, caption = 'Cheat sheet for logical operations in R')
```

## 'print' function

This is a fuction normal in many programming languages that displays your answer from now on I'll be using this function to display values
```{r}
print(2)

print("this is a string")

print(2*3)

print(2 == 4/2)
```

## Variables
You can store values inside variables using "=" to use later on your code  

```{r}
a = 1
print(a)
```

You can also change its content by making modifications

```{r}
a = a + 1
print(a)
```

or by changing its value directly
```{r}
a = 'new value'
print(a)
```

# Vectors

A vector is a list of items

```{r}
a = c(1,2,3)
print(a)
```
# if ... else statements
Executes some code if the condition is met, execute other code otherwise 

```{r}
a = 42
b = 26

if (a > b) {
  print("a is greater than b")
} else if (a < b) {
  print ("a is less than b ")
} else{
  print ("a is equal to b")
}
```

You can also combine conditions using the symbols "&" or "|" the first one meaning 'and', the last one meaning 'or'

Using &
```{r}
a <- 2
b <- 1
c <- 3

if (a > b & c > a){
  print("Both conditions are true")
}
```

Using |

```{r}
a <- 2
b <- 1
c <- 3

if (a > b | c < a){
  print("At least one condition is true")
}
```

# Loops

loops are chunks of codes that keeps repeating itself until a condition is considered false

## Loop 'while'

This code runs while the condition is met

```{r}
variable = 0
while (variable < 10){
  variable = variable + 1
  print(variable)
}
```

## Loop 'for'

This code runs until 'variable' reaches the end of the your vector, in this case 3. Note that 'variable' changes its value every time the code is executed
```{r}
variable = 0
for (variable in c(1,2,3)){
  print(variable)
}
```

